---
title: "ToothGrowth"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
```

## Summary of the Data in the ToothGrowth Dataset

The ToothGrowth dataset in R contains measures of the effect of various dose levels of Vitamin C (0.5, 1, and 2 mg/day) on tooth growth in 60 guinea pigs. For each dose level, 2 different delivery methods are compared orage juice (OJ) and ascorbic acid (VC).

```{r}
head(ToothGrowth)

summary(ToothGrowth)
```

A number of observations can be made regarding the effect of Vitamin C on tooth gowth.

Higher dose levels of Vitamin C seems to yield to greater tooth gowth.
That seems to be the cases regardless of the delivery method.

```{r warning=FALSE}
ggplot(data = ToothGrowth, aes(dose, len, color = supp)) + geom_point() + geom_smooth()
```

Orange juice seems to be a more efficient delivery method than ascorbic acid for dose levels 0.5 and 1 mg/day. 
For 2mg/day it is not clear which delivery method is the most efficient.

```{r}
ggplot(data = ToothGrowth, aes(x=factor(dose), y=len, fill=supp)) + geom_boxplot()
```

## Hypothesis Testing

### Impact of Dose Levels on Tooth Growth

```{r}
tg05 <- subset(ToothGrowth, dose == 0.5) 
tg1 <- subset(ToothGrowth, dose == 1)
tg2 <- subset(ToothGrowth, dose == 2)
```

### 0.5 vs 1 mg/day

Null Hypothesis: Increasing Dosage from 0.5mg to 1mg does not yield to greater tooth gowth

```{r}
t.test(tg1$len - tg05$len, paired = FALSE, var.equal = FALSE, alternative = "greater")$conf
```
The 95% confidence interval is above 0 we therefore reject the null hypothesis.

### 1 vs 2 mg/day

Null Hypothesis: Increasing Dosage from 1mg to 2mg does not yield to greater tooth gowth

```{r}
t.test(tg2$len - tg1$len, paired = FALSE, var.equal = FALSE, alternative = "greater")$conf
```
The 95% confidence interval is above 0 we therefore reject the null hypothesis.

### Impact of Delivery Methods on Tooth Growth

```{r}
tgOJ <- subset(ToothGrowth, supp == "OJ") 
tgVC <- subset(ToothGrowth, supp == "VC")
```

Null Hypothesis: The delivery method has no ipact on tooth growth

```{r}
t.test(tgOJ$len - tgVC$len, paired = FALSE, var.equal = FALSE)$conf
```
The 95% confidence interval does not contain 0. We therefore reject the null hypothesis.

### Impact of Delivery Methods on Tooth Growth for Various Dose Levels

#### Dosage = 0.5mg/day

Null Hypothesis: For a dosage of 0.5mg/day, the delivery method has no impact on tooth growth

```{r}
tgOJ05 <- subset(tgOJ, dose == 0.5) 
tgVC05 <- subset(tgVC, dose == 0.5)

t.test(tgOJ05$len - tgVC05$len, paired = FALSE, var.equal = FALSE)$conf
```
The 95% confidence interval does not contain 0. We therefore reject the null hypothesis.


#### Dosage = 1mg/day

Null Hypothesis: For a dosage of 1mg/day, the delivery method has no impact on tooth growth

```{r}
tgOJ1 <- subset(tgOJ, dose == 1) 
tgVC1 <- subset(tgVC, dose == 1)

t.test(tgOJ1$len - tgVC1$len, paired = FALSE, var.equal = FALSE)$conf
```
The 95% confidence interval does not contain 0. We therefore reject the null hypothesis.

#### Dosage = 2mg/day

Null Hypothesis: For a dosage of 1mg/day, the delivery method has no impact on tooth growth

```{r}
tgOJ2 <- subset(tgOJ, dose == 2) 
tgVC2 <- subset(tgVC, dose == 2)

t.test(tgOJ2$len - tgVC2$len, paired = FALSE, var.equal = FALSE)$conf
```
The 95% confidence interval contains 0. We therefore fail to reject the null hypothesis.

## Conclusions

* Higher dose levels of Vitamin C yield to greater tooth gowth.
* Orange juice a more efficient delivery method than ascorbic acid for dose levels of 0.5mg/day and 1mg/day
* The delivery methd does not seem to impact tooth gowth for dose levels of 2mg/day

## Assumptions

One of the assumptions made in this study is that the variance between different dose levels and delivery methods are not equal.

